 Title:    01_format_extracted_literature_data.Rmd
 Project:  ##BHI - Brain Health Initiative
 Purpose : To align IHME age groups to literature data age groups
           
 Author: USERNAME
 Date: 2024 06 03

Setting up environment
```{r}
# clear environment
rm(list = ls())

# Set up directory roots #
if (!exists("code_repo"))  {
  code_repo <- unname(ifelse(Sys.info()['sysname'] == "Windows", "ADDRESS", paste0("ADDRESS", Sys.info()['user'][1], "ADDRESS")))
}

# packages
library(data.table)
library(ggplot2)
library(tidyverse)

# file paths
file_dir <- 'FILEPATH'
data_output_dir <- paste0(file_dir, "FILEPATH")
condition_dt_path <- paste0(file_dir, "FILEPATH")
age_group_id_path <- paste0(file_dir, "FILEPATH")
toc_renaming_table_path <- paste0(file_dir, "FILEPATH")
direct_spending_path <- paste0(file_dir, "FILEPATH")

# functions
source("FILEPATH")
source(file.path("FILEPATH"))
source(paste0(code_repo, 'FILEPATH'))

```

Hierarchy and data info
```{r}
release_num <- 16

# get hierarchy
hierarchy <- get_location_metadata(
  location_set_id = 35,
  release_id = release_num
)

# storing country level location ids
lvl3_locs <- hierarchy[level == 3]$location_id

# storing condition name 
condition_name <- "Stroke"

# printing condition name
print(condition_name)

```

Load data
```{r}
# condition data
condition_dt <- fread(condition_dt_path)
unique(condition_dt$cause_name)

# making sure value column = continuous/numeric
condition_dt[, value := as.numeric(value)]

# removing empty rows
condition_dt <- condition_dt[!location_name == ""]

# adding spending type column
condition_dt[, spending_type := `health_function/spending category`]

# formatting gbd sex id
condition_dt[, gbd_sex_id := as.integer(gbd_sex_id)]

# age group id table
age_group_id_table <- fread(age_group_id_path)

# direct spending data
direct_spending_dt <- fread(direct_spending_path)

# toc renaming table
toc_renaming_table <- fread(toc_renaming_table_path)

```

Convert direct spending dt from wide to long
```{r}
direct_spending_dt <- melt(direct_spending_dt, id.vars = c("location_id",
                                                   "location_name",
                                                   "year_id",
                                                   "age_group_id",
                                                   "age_group_name",
                                                   "sex_id",
                                                   "cause_id",
                                                   "cause_name",
                                                   "incident_prevalent_cases"), variable.name = "new_spending_type", value.name = "bhi_value",
                   measure.vars = c("AM", "ED", "HH", "IP", "NF", "RX"))

# adding figure spending type column
direct_spending_dt[, type_of_care := ifelse(new_spending_type == "AM", "raked_mean_AM",
                                                   ifelse(new_spending_type == "ED", "raked_mean_ED",
                                                          ifelse(new_spending_type == "HH", "raked_mean_HH",
                                                                 ifelse(new_spending_type == "IP", "raked_mean_IP",
                                                                        ifelse(new_spending_type == "NF", "raked_mean_NF",
                                                                               ifelse(new_spending_type == "RX", "raked_mean_RX", NA))))))]

# removing new spending type column
direct_spending_dt[, new_spending_type := NULL]
```

Data cleaning
```{r}
# data info
# age groups
condition_age_groups <- sort(unique(condition_dt$population_age))
print(condition_age_groups)

# locations
condition_locs <- sort(unique(condition_dt$location_name))
print(condition_locs)

# renaming location names to match hierarchy
condition_dt[location_name == "United States", location_name := hierarchy[location_id == 102]$location_name]
condition_dt[location_name == "USA", location_name := hierarchy[location_id == 102]$location_name]
condition_dt[location_name == "Czech Republic", location_name := hierarchy[location_id == 47]$location_name]
condition_dt[location_name == "Russia", location_name := hierarchy[location_id == 62]$location_name]
condition_dt[location_name == "Sao Paolo", location_name := hierarchy[location_id == 4775]$location_name]
condition_dt[location_name == "Taiwan", location_name := hierarchy[location_id == 8]$location_name]
condition_dt[location_name == "Turkey", location_name := hierarchy[location_id == 155]$location_name]
condition_dt[location_name == "South Korea", hierarchy[location_id == 68]$location_name]
condition_dt[location_name == "United Kindgom", hierarchy[location_id == 95]$location_name]

# removing non-country-specific rows
condition_dt <- condition_dt[!location_name == "Europe"]

# updating population ages
condition_dt[population_age == "0 to under 2", population_age := "<2"]
condition_dt[population_age == "0 to 14", population_age := "0-14"]
condition_dt[population_age == "0 to 19", population_age := "0-19"]
condition_dt[population_age == "3 months to 4 years", population_age := "0-4"]
condition_dt[population_age == "91 days to 14 years", population_age := "0-14"]
condition_dt[population_age == "0 to 44", population_age := "0-44"]
condition_dt[population_age == "2 to under 5", population_age := "2-4"]
condition_dt[population_age == "2 to 30", population_age := "2-30"]
condition_dt[population_age == "3 to 17", population_age := "3-17"]
condition_dt[population_age == "5 to 9", population_age := "5-9"]
condition_dt[population_age == "5 to 11", population_age := "5-11"]
condition_dt[population_age == "5 to 14", population_age := "5-14"]
condition_dt[population_age == "5 to 17", population_age := "5-17"]
condition_dt[population_age == "5 to 91", population_age := "5-91"]
condition_dt[population_age == "6 to 18", population_age := "6-18"]
condition_dt[population_age == "7 to 18", population_age := "7-18"]
condition_dt[population_age == "8 to 17", population_age := "8-17"]
condition_dt[population_age == "10 to 14", population_age := "10-14"]
condition_dt[population_age == "11 to 14", population_age := "11-14"]
condition_dt[population_age == "11 to 17", population_age := "11-17"]
condition_dt[population_age == "11 to 18", population_age := "11-18"]
condition_dt[population_age == "12 to 17", population_age := "12-17"]
condition_dt[population_age == "12 to 18", population_age := "12-18"]
condition_dt[population_age == "15 to 17", population_age := "15-17"]
condition_dt[population_age == "15 years to 19 years", population_age := "15-19"]
condition_dt[population_age == "15 to 20", population_age := "15-20"]
condition_dt[population_age == "16 to 64", population_age := "16-64"]
condition_dt[population_age == "16 to 75", population_age := "16-75"]
condition_dt[population_age == "16 to 85", population_age := "16-85"]
condition_dt[population_age == "18 to 34", population_age := "18-34"]
condition_dt[population_age == "18 to 54", population_age := "18-54"]
condition_dt[population_age == "18 to 60", population_age := "18-60"]
condition_dt[population_age == "18 to 64", population_age := "18-64"]
condition_dt[population_age == "18 to 65", population_age := "18-65"]
condition_dt[population_age == "18 to 70", population_age := "18-70"]
condition_dt[population_age == "18 to 75", population_age := "18-75"]
condition_dt[population_age == "19 to 65", population_age := "19-65"]
condition_dt[population_age == "20 to 29", population_age := "20-29"]
condition_dt[population_age == "20 to 68", population_age := "20-68"]
condition_dt[population_age == "30 to 39", population_age := "30-39"]
condition_dt[population_age == "31 to 80", population_age := "31-80"]
condition_dt[population_age == "35 to 49", population_age := "35-49"]
condition_dt[population_age == "40 to 49", population_age := "40-49"]
condition_dt[population_age == "41 to 64", population_age := "41-64"]
condition_dt[population_age == "45 to 64", population_age := "45-64"]
condition_dt[population_age == "50 to 59", population_age := "50-59"]
condition_dt[population_age == "50 to 64", population_age := "50-64"]
condition_dt[population_age == "60 to 69", population_age := "60-69"]
condition_dt[population_age == "65 to 79", population_age := "65-79"]
condition_dt[population_age == "65 to 74", population_age := "65-74"]
condition_dt[population_age == "65 to 75", population_age := "65-75"]
condition_dt[population_age == "70 to 79", population_age := "70-79"]
condition_dt[population_age == "74 to 88", population_age := "74-88"]
condition_dt[population_age == "75 to 84", population_age := "75-84"]
condition_dt[population_age == "79 to 96", population_age := "79-96"]
condition_dt[population_age == "under 18", population_age := "<=17"]
condition_dt[population_age == "65 plus", population_age := "65+"]
condition_dt[population_age == "all", population_age := "all ages"]

```

Converting all literature values to annual values
```{r}

# proportion of year for each sampling period
condition_dt[sampling_period == "1 day", proportion_of_year := (1/365)]
condition_dt[sampling_period == "day", proportion_of_year := (1/365)]
condition_dt[sampling_period == "week", proportion_of_year := (1/52)]
condition_dt[sampling_period == "15 days", proportion_of_year := (3/73)]
condition_dt[sampling_period == "1.5 months (6 weeks)", proportion_of_year := (6/52)]
condition_dt[sampling_period == "1 month", proportion_of_year := (1/12)]
condition_dt[sampling_period == "month", proportion_of_year := (1/12)]
condition_dt[sampling_period == "3 months", proportion_of_year := (3/12)]
condition_dt[sampling_period == "6 months", proportion_of_year := (6/12)]
condition_dt[sampling_period == "9 months", proportion_of_year := (9/12)]
condition_dt[sampling_period == "10 months", proportion_of_year := (10/12)]
condition_dt[sampling_period == "12 months", proportion_of_year := (12/12)]
condition_dt[sampling_period == "year", proportion_of_year := (12/12)]
condition_dt[sampling_period == "1 year", proportion_of_year := (12/12)]
condition_dt[sampling_period == "18 months", proportion_of_year := (18/12)]
condition_dt[sampling_period == "24 months", proportion_of_year := (24/12)]
condition_dt[sampling_period == "36 months", proportion_of_year := (36/12)]
condition_dt[sampling_period == "48 months", proportion_of_year := (48/12)]
condition_dt[sampling_period == "5 years", proportion_of_year := (60/12)]

# calculate annual value based on proportion of year
condition_dt[, annual_value := value / proportion_of_year]

```

Consolidating spending types and renaming data year to year id
```{r}
condition_dt <- merge(condition_dt, toc_renaming_table, by = "spending_type", all.x = TRUE)
setnames(condition_dt, old = "data_year", new = "year_id")

```

Only retaining direct spending types of care
```{r}
ds_toc <- c("raked_mean_AM",
            "raked_mean_ED",
            "raked_mean_HH",
            "raked_mean_IP",
            "raked_mean_NF",
            "raked_mean_RX")

condition_dt <- condition_dt[type_of_care %in% ds_toc & location_id %in% lvl3_locs & year_id %in% c(2000:2019)]

```

Aligning cause names between datasets
```{r}
# updating cause name in data to match direct spending pre-adjusted data
cause_table <- unique(direct_spending_dt[, .(cause_id, cause_name)])
condition_dt[, cause_name := NULL]
condition_dt <- merge(condition_dt, cause_table, by = "cause_id", all.x = T)

```

Storing population age group ids
```{r}
# all age groups
gbd_age_0_4 <- c(1)
gbd_age_0_14 <- c(1, 6:7)
gbd_age_0_44 <- c(1, 6:13)
gbd_age_2_4 <- c(1)
gbd_age_2_30 <- c(1, 6:11)
gbd_age_3_17 <- c(1, 6:8) 
gbd_age_5_9 <- c(6)
gbd_age_5_11 <- c(6:7)
gbd_age_5_14 <- c(6:7)
gbd_age_5_17 <- c(6:8)
gbd_age_5_91 <- c(6:20, 30, 160)
gbd_age_6_18 <- c(6:8)
gbd_age_7_18 <- c(6:8)
gbd_age_8_17 <- c(6:8)
gbd_age_10_14 <- c(7)
gbd_age_11_14 <- c(7)
gbd_age_11_17 <- c(7:8)
gbd_age_11_18 <- c(7:8)
gbd_age_12_17 <- c(7:8)
gbd_age_12_18 <- c(7:8)
gbd_age_15_17 <- c(8)
gbd_age_15_18 <- c(8)
gbd_age_15_19 <- c(8)
gbd_age_15_20 <- c(8:9)
gbd_age_15_60 <- c(8:16)
gbd_age_15_64 <- c(8:17)
gbd_age_16_48 <- c(8:14)
gbd_age_16_64 <- c(8:17)
gbd_age_16_74 <- c(8:19)
gbd_age_16_75 <- c(8:20)
gbd_age_16_85 <- c(8:20, 30, 160)
gbd_age_18_30 <- c(8:10)
gbd_age_18_34 <- c(8:11)
gbd_age_18_54 <- c(8:15)
gbd_age_18_60 <- c(8:17)
gbd_age_18_64 <- c(8:17)
gbd_age_18_65 <- c(8:18)
gbd_age_18_70 <- c(8:19)
gbd_age_18_75 <- c(8:20)
gbd_age_18_80 <- c(8:20)
gbd_age_19_65 <- c(8:18)
gbd_age_19_78 <- c(8:20)
gbd_age_19_85 <- c(8:20, 30, 160)
gbd_age_20_60 <- c(9:17)
gbd_age_20_64 <- c(9:17)
gbd_age_20_68 <- c(9:18)
gbd_age_24_76 <- c(9:20)
gbd_age_24_77 <- c(9:20)
gbd_age_35_49 <- c(12:14)
gbd_age_45_64 <- c(14:17)
gbd_age_50_64 <- c(15:17)
gbd_age_65_74 <- c(18:19)
gbd_age_65_75 <- c(18:20)
gbd_age_65_79 <- c(18:20)
gbd_age_74_88 <- c(19:20, 30, 160)
gbd_age_75_84 <- c(20, 30)
gbd_age_79_96 <- c(20, 30, 160)
gbd_age_85_100 <- c(160)

# plus
gbd_age_5_plus <- c(6:20, 30, 160)
gbd_age_15_plus <- c(8:20, 30, 160)
gbd_age_17_plus <- c(8:20, 30, 160)
gbd_age_18_plus <- c(8:20, 30, 160)
gbd_age_19_plus <- c(8:20, 30, 160)
gbd_age_20_plus <- c(9:20, 30, 160)
gbd_age_25_plus <- c(10:20, 30, 160)
gbd_age_30_plus <- c(11:20, 30, 160)
gbd_age_35_plus <- c(12:20, 30, 160)
gbd_age_40_plus <- c(13:20, 30, 160)
gbd_age_45_plus <- c(14:20, 30, 160)
gbd_age_50_plus <- c(15:20, 30, 160)
gbd_age_55_plus <- c(16:20, 30, 160)
gbd_age_60_plus <- c(17:20, 30, 160)
gbd_age_65_plus <- c(18:20, 30, 160)
gbd_age_70_plus <- c(19:20, 30, 160)
gbd_age_75_plus <- c(20, 30, 160)
gbd_age_80_plus <- c(30, 160)
gbd_age_85_plus <- c(160)

# age and under
gbd_age_17_and_under <- c(1, 6:8)
gbd_age_18_and_under <- c(1, 6:8)

# under
gbd_age_under_1 <- c(1)
gbd_age_under_2 <- c(1) 
gbd_age_under_5 <- c(1)
gbd_age_under_18 <- c(1, 6:8)
gbd_age_under_50 <- c(1, 6:14)
gbd_age_under_65 <- c(1, 6:17)

# over 60
gbd_age_over_60 <- c(17:20, 30, 160)

# all ages
gbd_age_all_ages <- c(1, 6:20, 30, 160)

```

Aggregating BHI direct spending data to match lit population age groups
```{r}
# storing titles
titles <- sort(unique(condition_dt$title))

# bhi cause id 
bhi_cause_id <- unique(condition_dt$cause_id)
print(bhi_cause_id)

# creating empty list to store data
my_list <- list()

# titles for loop
for (t in titles) {
  
  temp_title <- condition_dt[title == t]
  
  
  locs <- sort(unique(temp_title$location_name))
  
  # location for loop
  for (loc in locs){
    
    temp_loc <- temp_title[location_name == loc]  
    years <- sort(unique(temp_loc$year_id))
    
    # year for loop
    for (year in years){
      
      temp_year <- temp_loc[year_id == year]
      temp_age_group_ids <- sort(unique(temp_year$population_age))
      
      # age group ids for loop
      for (age in temp_age_group_ids) {
        
        temp_ages <- temp_year[population_age == age]
        gbd_sex_ids <- sort(unique(temp_ages$gbd_sex_id))
        
        # sex for loop
        for (sex in gbd_sex_ids){
          
          # updating sex
          sex_val <- sex
          temp_gbd_sex <- temp_ages[gbd_sex_id == sex_val]
          temp_spending_types <- sort(unique(temp_gbd_sex$type_of_care))
          
          # filtering for each age group
            temp_age_group_id_dt <- copy(temp_gbd_sex)
            
            # start of age group id for loop
            if(age == "<2"){
              temp_age_group_id_vals <- gbd_age_under_2
            } else if(age == "0-4"){
              temp_age_group_id_vals <- gbd_age_0_4
            } else if(age == "0-14"){
              temp_age_group_id_vals <- gbd_age_0_14
            } else if(age == "0-44"){
              temp_age_group_id_vals <- gbd_age_0_44
            } else if(age == "2-4"){
              temp_age_group_id_vals <- gbd_age_2_4
            } else if(age == "2-30"){
              temp_age_group_id_vals <- gbd_age_2_30
            } else if(age == "3-17"){
              temp_age_group_id_vals <- gbd_age_3_17
            } else if(age == "5-9"){
              temp_age_group_id_vals <- gbd_age_5_9
            } else if(age == "5-11"){
              temp_age_group_id_vals <- gbd_age_5_11
            } else if(age == "5-14"){
              temp_age_group_id_vals <- gbd_age_5_14
            } else if(age == "5-17"){
              temp_age_group_id_vals <- gbd_age_5_17
            } else if(age == "5-91"){
              temp_age_group_id_vals <- gbd_age_5_91
            } else if(age == "6-18"){
              temp_age_group_id_vals <- gbd_age_6_18
            } else if(age == "7-18"){
              temp_age_group_id_vals <- gbd_age_7_18
            } else if(age == "8-17"){
              temp_age_group_id_vals <- gbd_age_8_17
            } else if(age == "10-14"){
              temp_age_group_id_vals <- gbd_age_10_14
            } else if(age == "11-14"){
              temp_age_group_id_vals <- gbd_age_11_14
            } else if(age == "11-17"){
              temp_age_group_id_vals <- gbd_age_11_17
            } else if(age == "11-18"){
              temp_age_group_id_vals <- gbd_age_11_18
            } else if(age == "12-17"){
              temp_age_group_id_vals <- gbd_age_12_17
            } else if(age == "12-18"){
              temp_age_group_id_vals <- gbd_age_12_18
            } else if(age == "15-17"){
              temp_age_group_id_vals <- gbd_age_15_17
            } else if(age == "15-18"){
              temp_age_group_id_vals <- gbd_age_15_18
            } else if(age == "15-19"){
              temp_age_group_id_vals <- gbd_age_15_19
            } else if(age == "15-20"){
              temp_age_group_id_vals <- gbd_age_15_20
            } else if(age == "15-60"){
              temp_age_group_id_vals <- gbd_age_15_60
            } else if(age == "15-64"){
              temp_age_group_id_vals <- gbd_age_15_64
            } else if(age == "16-48"){
              temp_age_group_id_vals <- gbd_age_16_48
            } else if(age == "16-64"){
              temp_age_group_id_vals <- gbd_age_16_64
            } else if(age == "16-74"){
              temp_age_group_id_vals <- gbd_age_16_74
            } else if(age == "16-75"){
              temp_age_group_id_vals <- gbd_age_16_75
            } else if(age == "16-85"){
              temp_age_group_id_vals <- gbd_age_16_85
            } else if(age == "18-30"){
              temp_age_group_id_vals <- gbd_age_18_30
            } else if(age == "18-34"){
              temp_age_group_id_vals <- gbd_age_18_34
            } else if(age == "18-54"){
              temp_age_group_id_vals <- gbd_age_18_54
            } else if(age == "18-60"){
              temp_age_group_id_vals <- gbd_age_18_60
            } else if(age == "18-64"){
              temp_age_group_id_vals <- gbd_age_18_64
            } else if(age == "18-65"){
              temp_age_group_id_vals <- gbd_age_18_65
            } else if(age == "18-70"){
              temp_age_group_id_vals <- gbd_age_18_70
            } else if(age == "18-75"){
              temp_age_group_id_vals <- gbd_age_18_75
            } else if(age == "18-80"){
              temp_age_group_id_vals <- gbd_age_18_80
            } else if(age == "19-65"){
              temp_age_group_id_vals <- gbd_age_19_65
            } else if(age == "19-78"){
              temp_age_group_id_vals <- gbd_age_19_78
            } else if (age == "19-85"){
              temp_age_group_id_vals <- gbd_age_19_85
            } else if (age == "20-60"){
              temp_age_group_id_vals <- gbd_age_20_60
            } else if (age == "20-64"){
              temp_age_group_id_vals <- gbd_age_20_64
            } else if (age == "20-68"){
              temp_age_group_id_vals <- gbd_age_20_68
            } else if (age == "24-76"){
              temp_age_group_id_vals <- gbd_age_24_76
            } else if (age == "24-77"){
              temp_age_group_id_vals <- gbd_age_24_77
            } else if (age == "35-49"){
              temp_age_group_id_vals <- gbd_age_35_49
            } else if (age == "45-64"){
              temp_age_group_id_vals <- gbd_age_45_64
            } else if (age == "50-64"){
              temp_age_group_id_vals <- gbd_age_50_64
            } else if (age == "65-74"){
              temp_age_group_id_vals <- gbd_age_65_74
            } else if (age == "65-75"){
              temp_age_group_id_vals <- gbd_age_65_75
            } else if (age == "65-79"){
              temp_age_group_id_vals <- gbd_age_65_79
            } else if (age == "74-88"){
              temp_age_group_id_vals <- gbd_age_74_88
            } else if (age == "75-84"){
              temp_age_group_id_vals <- gbd_age_75_84
            } else if (age == "79-96"){
              temp_age_group_id_vals <- gbd_age_79_96
            } else if (age == "85-100"){
              temp_age_group_id_vals <- gbd_age_85_100
            } else if (age == "5+"){
              temp_age_group_id_vals <- gbd_age_5_plus
            } else if (age == "15+"){
              temp_age_group_id_vals <- gbd_age_15_plus
            } else if (age == "17+"){
              temp_age_group_id_vals <- gbd_age_17_plus
            } else if (age == "18+"){
              temp_age_group_id_vals <- gbd_age_18_plus
            } else if (age == "19+"){
              temp_age_group_id_vals <- gbd_age_19_plus
            } else if (age == ">18"){
              temp_age_group_id_vals <- gbd_age_19_plus
            } else if (age == "20+"){
              temp_age_group_id_vals <- gbd_age_20_plus
            } else if (age == "25+"){
              temp_age_group_id_vals <- gbd_age_25_plus
            } else if (age == "30+"){
              temp_age_group_id_vals <- gbd_age_30_plus
            } else if (age == "35+"){
              temp_age_group_id_vals <- gbd_age_35_plus
            } else if (age == "40+"){
              temp_age_group_id_vals <- gbd_age_40_plus
            } else if (age == "45+"){
              temp_age_group_id_vals <- gbd_age_45_plus
            } else if (age == "50+"){
              temp_age_group_id_vals <- gbd_age_50_plus
            } else if (age == "55+"){
              temp_age_group_id_vals <- gbd_age_55_plus
            } else if (age == "60+"){
              temp_age_group_id_vals <- gbd_age_60_plus
            } else if (age == "65+"){
              temp_age_group_id_vals <- gbd_age_65_plus
            } else if (age == "70+"){
              temp_age_group_id_vals <- gbd_age_70_plus
            } else if (age == "75+"){
              temp_age_group_id_vals <- gbd_age_75_plus
            } else if (age == "80+"){
              temp_age_group_id_vals <- gbd_age_80_plus
            } else if (age == "85+"){
              temp_age_group_id_vals <- gbd_age_85_plus
            } else if (age == "<=17"){
              temp_age_group_id_vals <- gbd_age_17_and_under
            } else if (age == "<=18"){
              temp_age_group_id_vals <- gbd_age_18_and_under
            } else if (age == "under 1"){
              temp_age_group_id_vals <- gbd_age_under_1
            } else if (age == "under 5"){
              temp_age_group_id_vals <- gbd_age_under_5
            } else if (age == "under 50"){
              temp_age_group_id_vals <- gbd_age_under_50
            } else if (age == "under 65"){
              temp_age_group_id_vals <- gbd_age_under_65
            } else if (age == "over 60"){
              temp_age_group_id_vals <- gbd_age_over_60
            } else if (age == "all ages"){
              temp_age_group_id_vals <- gbd_age_all_ages
            } else {
              temp_age_group_id_vals <- is.numeric(NA)
            }
             temp_loc_id <- unique(temp_age_group_id_dt$location_id)
             
             # storing sex ids in environment
             if(sex_val == 3){
               temp_sex_id <- c(1, 2)
             } else if (sex_val == 1){ 
               temp_sex_id <- 1
             } else{
               temp_sex_id <- 2
             }
             
             # creating temp dataset
              temp_bhi_ds_dt <- direct_spending_dt[location_id %in% temp_loc_id & year_id %in% year & age_group_id %in% temp_age_group_id_vals & sex_id %in% temp_sex_id & cause_id %in% bhi_cause_id & type_of_care %in% temp_spending_types]
              
              # if else statement for dts with only 1 row of data
              if((length(unique(temp_bhi_ds_dt$age_group_id)) + sex_val) > 3){
              
                # aggregate bhi ds values and prev values at same time
                temp_bhi_ds_dt_agg <- data.table(aggregate(cbind(bhi_value, incident_prevalent_cases) ~ location_id + year_id + sex_id + cause_id + cause_name + type_of_care, data = temp_bhi_ds_dt, FUN = sum))
                
                # if state to aggregate over sex if gbd sex id = 3
                if(sex == 3){
                  temp_bhi_ds_dt_agg <- data.table(aggregate(cbind(bhi_value, incident_prevalent_cases) ~ location_id + year_id + cause_id + cause_name + type_of_care, data = temp_bhi_ds_dt_agg, FUN = sum))
                  temp_bhi_ds_dt_agg[, sex_id := 3]
                } else{
                  # keep sex id
                } 
                  } else {
                    
                  temp_bhi_ds_dt_agg <- copy(temp_bhi_ds_dt)
                }
              
              # creating sex id column
              temp_age_group_id_dt[, sex_id := sex_val]
              
              # creating population age column in dataset
              temp_bhi_ds_dt_agg[, population_age := age]
              
              # merging bhi values with age group id dt
              temp_age_group_id_dt <- merge(temp_age_group_id_dt, temp_bhi_ds_dt_agg, by = c("location_id", "year_id", "sex_id", "cause_id", "cause_name", "type_of_care", "population_age"), all.x = T)
              
              # creating unique list name and storing in list
              list_name <- paste0(c(t, loc, year, age, sex), collapse = "_")
              my_list[[list_name]] <- copy(temp_age_group_id_dt)
          
        } # end of gbd sex id loop
        
      }   # end of age group id loop

    } # end of year loop
    
    
  } # end of location loop
  
  
} # end of title loop

# binding list into one dt
dt <- rbindlist(my_list)

```  

Cleaning up aggregation level column
```{r}
# reorder dataset
setorder(dt, title, location_id, year_id, population_age, sex_id)

# reformatting aggregation level column
dt[aggregation_level == "cumulative value", aggregation_level := "cumulative"]
dt[aggregation_level == "cumulative_value", aggregation_level := "cumulative"]
dt[aggregation_level == "cumulative values", aggregation_level := "cumulative"]
dt[aggregation_level == "total", aggregation_level := "cumulative"]
dt[aggregation_level == "per patient values", aggregation_level := "per patient"]
dt[aggregation_level == "unit", aggregation_level := "unit cost"]

# calculating final value if multiplication factor present
dt[, value_final := annual_value * mult_factor]
summary(dt$value_final)

```

Creating per prevalence values
```{r}
# create separate prevalence value columns
dt[, bhi_inc_prev_val := incident_prevalent_cases]
setnames(dt, old = "incident_prevalent_cases", new = "lit_inc_prev_val")

# pre-adjusted values set to 1 for non cumulative values
dt[!aggregation_level == "cumulative", lit_inc_prev_val := 1]

# calculating per prevalent case values
dt[, `:=` (bhi_per_inc_prev_val = (bhi_value / bhi_inc_prev_val), lit_per_inc_prev_val = (value_final / lit_inc_prev_val))]
```

Saving out dt
```{r}
fwrite(dt, paste0(data_output_dir, condition_name, "FILEPATH"))

```
