 Title:    02_consolidate_literature_datasets.Rmd
 Project:  ##BHI - Brain Health Initiative
 Purpose : To consolidate all the reallocated values datasets for the literature review
           
 Author: USERNAME
 Date: 2025 03 07

Setting up environment
```{r}
# clear environment
rm(list = ls())

# Set up directory roots #
if (!exists("code_repo"))  {
  code_repo <- unname(ifelse(Sys.info()['sysname'] == "Windows", "USERNAME", paste0("USERNAME", Sys.info()['user'][1], "USERNAME")))
}

# packages
library(data.table)
library(ggplot2)

# file paths
data_dir <- "FILEPATH"
data_output_dir <- paste0(data_dir, "FILEPATH")
figure_output_dir <- "FILEPATH"

adhd_path <- paste0(data_dir, "FILEPATH")
alzheimers_path <- paste0(data_dir, "FILEPATH")
anxiety_path <- paste0(data_dir, "FILEPATH")
autism_path <- paste0(data_dir, "FILEPATH")
bipolar_path <- paste0(data_dir, "FILEPATH")
brain_cancer_path <- paste0(data_dir, "FILEPATH")
depression_path <- paste0(data_dir, "FILEPATH")
eating_path <- paste0(data_dir, "FILEPATH")
epilepsy_path <- paste0(data_dir, "FILEPATH")
headache_disorders_path <- paste0(data_dir, "FILEPATH")
meningitis_path <- paste0(data_dir, "FILEPATH")
ms_path <- paste0(data_dir, "FILEPATH")
parkinsons_path <- paste0(data_dir, "FILEPATH")
schizophrenia_path <- paste0(data_dir, "FILEPATH")
stroke_path <- paste0(data_dir, "FILEPATH")

# functions
source("FILEPATH")
source(file.path("FILEPATH"))
source(paste0(code_repo, 'FILEPATH'))

```

location and date info
```{r}
# save date
date <- gsub('-', '_', Sys.Date())

gbd_release_id <- 16

# get hierarchy
hierarchy <- get_location_metadata(
  location_set_id = 35,
  release_id = gbd_release_id
)

```

Columns to keep from all datasets
```{r}
# columns to keep 
keep_cols <- c("cause_name", "cause_id", "title", "location_name", 
               "location_id", "year_id", "sex_id", "type_of_care", "population_age", 
               "spending_type", "aggregation_level", "bhi_component", "mult_factor", "sample_size", 
               "value", "value_final", "lit_inc_prev_val", "lit_per_inc_prev_val", "bhi_value", 
               "bhi_inc_prev_val", "bhi_per_inc_prev_val", "unit", "sampling_period", "base_year", 
               "currency")

```

Reading in datasets with desired columns
```{r}
# read in datasets
adhd_dt <- fread(adhd_path)[, ..keep_cols]
alzheimers_dt <- fread(alzheimers_path)[, ..keep_cols]
anxiety_dt <- fread(anxiety_path)[, ..keep_cols]
autism_dt <- fread(autism_path)[, ..keep_cols]
bipolar_dt <- fread(bipolar_path)[, ..keep_cols]
brain_cancer_dt <- fread(brain_cancer_path)[, ..keep_cols]
depression_dt <- fread(depression_path)[, ..keep_cols]
eating_dt <- fread(eating_path)[, ..keep_cols]
epilepsy_dt <- fread(epilepsy_path)[, ..keep_cols]
headache_disorders_dt <- fread(headache_disorders_path)[, ..keep_cols]
meningitis_dt <- fread(meningitis_path)[, ..keep_cols]
ms_dt <- fread(ms_path)[, ..keep_cols]
parkinsons_dt <- fread(parkinsons_path)[, ..keep_cols]
schizophrenia_dt <- fread(schizophrenia_path)[, ..keep_cols]
stroke_dt <- fread(stroke_path)[, ..keep_cols]

# bind list
dt <- rbindlist(list(adhd_dt, anxiety_dt, autism_dt, alzheimers_dt, bipolar_dt, 
                     brain_cancer_dt, depression_dt, eating_dt, epilepsy_dt, headache_disorders_dt, 
                     meningitis_dt, ms_dt, parkinsons_dt, schizophrenia_dt, stroke_dt))

```

Currency conversion
```{r}
# adding ihme loc id
dt <- merge(dt, hierarchy[, .(location_id, ihme_loc_id)], by = "location_id", all.x = T)

# updating currency column
dt[, currency_col := ifelse(grepl('Euro', currency, ignore.case = T), 'EUR',
                                     ifelse(grepl("EUR", currency), 'EUR',
                                          ifelse(grepl("US", currency), 'USD',
                                                  ifelse(grepl('USD', currency, ignore.case = T), 'USD', 'LCU'))))]

## currency conversion
dt <- currency_conversion(dt,
                          col.loc = 'ihme_loc_id',
                          col.value = 'lit_per_inc_prev_val',
                          col.currency = 'currency_col',
                          col.currency.year = 'base_year',
                          base.unit = 'USD',
                          base.year = 2021,
                          converter.version = 7,
                          simplify = F)

```

Saving out the final dataset
```{r}
fwrite(dt, paste0(data_output_dir, "FILEPATH"))
```