Title:    03_generate_covariates.Rmd 
 Project:  ##BHI - Brain Health Initiative
 Purpose : Generate covariate values for the direct spending regression model and correlation matrix
           
     
 Author: USERNAME
 Date: 2025 03 07
 
Setting up environment
```{r}
# clear environment
rm(list = ls())

# Set up directory roots #
if (!exists("code_repo"))  {
  code_repo <- unname(ifelse(Sys.info()['sysname'] == "Windows", "FILEPATH", paste0("FILEPATH", Sys.info()['user'][1], "FILEPATH")))
}

# packages
library(data.table)
library(lme4)
library(lmerTest)

# file paths
data_dir <- "FILEPATH"
model_data_path <- paste0(data_dir, "FILEPATH")
domestic_health_path <- "FILEPATH"
fig_out_dir <-  "FILEPATH"

# functions
source("FILEPATH")
source("FILEPATH")
source('FILEPATH')
source("FILEPATH")
source(paste0(code_repo, 'FILEPATH'))

# release number
gbd_release_id <- 16

# get hierarchy
hierarchy <- get_location_metadata(
  location_set_id = 35,
  release_id = gbd_release_id
)

```

Import data
```{r}
# main data
dt <- fread(model_data_path)

# domestic health spending dataset
dhe_dt <- fread(domestic_health_path)

```

Data info
```{r}
locs <- sort(unique(dt$location_id))
years <- c(2000:2019)
age_groups <- c(1, 6:20, 30, 160)
sex <- c(1, 2)

```

################################################################################
COVARIATES
################################################################################

GDP PC
```{r}
# get gdp
gdp <- data.table(read_feather("FILEPATH"))
gdp <- gdp[scenario == 0 & year <= 2019 & year >= 2000]
gdp <- melt.data.table(gdp, id.vars = c("year", "iso3", "scenario"), value.name = "data_var", variable.name = "draw")
gdp <- gdp[, draw := str_replace(draw, "draw_", "")]
gdp <- gdp[, draw := as.numeric(as.character(draw))]
gdp[, scenario := NULL]

# condensing and renaming columns
gdp <- gdp[year %in% years,
           .(gdp_pc = mean(data_var)),
           by = .(year_id = year, ihme_loc_id = iso3)]

# creating gbd dataset
gdp <- merge(gdp[, .(ihme_loc_id, 
                        year_id, 
                        gdp_pc)], hierarchy[, .(location_id, 
                                                    location_name, 
                                                    ihme_loc_id)], by = 'ihme_loc_id')

# convert from 2022 USD to 2021 USD
gdp_dt_conv <- currency_conversion(gdp,
                           col.loc = 'ihme_loc_id',
                           col.value = 'gdp_pc',
                           currency = 'usd',
                           currency.year = 2022 ,
                           base.year = 2021,
                           base.unit = 'usd',
                           converter.version = 7,
                           simplify = F)

# storing columns to keep
keep_cols <- c("ihme_loc_id", "year_id", "location_id", "currency_year_new", "gdp_pc_new")

# condense dataset
gdp_dt <- gdp_dt_conv[, ..keep_cols]

# rename GDP per capita column
setnames(gdp_dt, old = "gdp_pc_new", new = "gdp_pc")

# removing extra columns
gdp_dt[, currency_year_new := NULL]

```

Fraction of population older than 65
```{r}
# pulling total population values
tot_pops_dt <- get_population(location_id = locs,
                       year_id = years,
                       age_group_id = 22,
                       sex_id = 3,
                       release_id = gbd_release_id
                       )

# adding ihme loc id 
tot_pops_dt <- merge(tot_pops_dt, hierarchy[, .(location_id, ihme_loc_id)], by = "location_id")

# condensing columns
tot_pops_dt <- tot_pops_dt[, .(ihme_loc_id, 
                               location_id,
                               year_id, 
                               population)]

#  get population values stratified by age
age_pop_dt <- get_population(location_id = locs,
                       year_id = years,
                       age_group_id = c(1, 6:20, 30:32, 235),
                       sex_id = sex,
                       release_id = gbd_release_id
                       )

# filtering for population values of 65+
pop_over65 <- age_pop_dt[age_group_id %in% c(18:20, 30:32, 235), .(location_id, year_id, age_group_id, sex_id, population)]

# aggregate population values by country and year for population totals of 65+ 
pop_over65_agg <- data.table(aggregate(population ~ location_id + year_id, data = pop_over65, FUN = sum))

# rename population agg col
setnames(pop_over65_agg, old = "population", new = "pop_65_and_over")

# merge with total population values
pop_dt <- merge(pop_over65_agg, tot_pops_dt, by = c("location_id", "year_id"), all.x = T)

# calculate fraction of population that is 65+
pop_dt[, pop_65_frac := pop_65_and_over / population]

# removing pop totes col
pop_dt[, `:=` (population = NULL, pop_65_and_over = NULL)]

```

Adding socio-demographic index (SDI)
```{r}
# retreiving sdi values
sdi_dt <- get_covariate_estimates(covariate_id = 881,
                                   location_id = locs,
                                   year_id =  years,
                                   release_id = gbd_release_id) 

# condense dataset
sdi_dt <- sdi_dt[, .(location_id, 
                    year_id,
                    sdi_val = mean_value)]

```

Adding health access and quality index (HAQI)
```{r}
# retreiving haqi values
haqi <- get_covariate_estimates(covariate_id = 1099,
                                location_id = locs,
                                year_id = years,
                                release_id = gbd_release_id)

# renaming haqi column
setnames(haqi, old = "mean_value", new = "haqi_val")

# condense dataset
haqi_dt <- haqi[, .(location_id, year_id, haqi_val)]

```

Adding domestic health spending
```{r}
# only retaining country-level locations
dhe_dt <- dhe_dt[location_id %in% locs]

```

Adding covariates to dataset
```{r}
dt <- merge(dt, gdp_dt, by = c("location_id", "ihme_loc_id", "year_id"), all.x = T)
dt <- merge(dt, pop_dt, by = c("location_id", "ihme_loc_id", "year_id"), all.x = T)
dt <- merge(dt, sdi_dt, by = c("location_id", "year_id"), all.x = T)
dt <- merge(dt, haqi_dt, by = c("location_id", "year_id"), all.x = T)
dt <- merge(dt, dhe_dt, by = c("location_id", "year_id"), all.x = T)

```

Adding sex label
```{r}
# adding sex label
dt[sex_id == 3, sex_label := "Both Sexes"]
dt[sex_id == 1, sex_label := "Male"]
dt[sex_id == 2, sex_label := "Female"]

```

Adding disorder classification
```{r}
# mental health disorders
mental_disorders <- c(559,
                      560,
                      562,
                      566,
                    	567,
                    	570,
                    	571,
                    	572,
                    	575,
                    	578,
                      579,
                      582,
                      585,
                      718,
                      9999)

# neurological health disorders
neuro_disorders <- c(332,
                     337,
                     477,
                     494,
                     543,
                     544,
                     545,
                     546,
                     554,
                     972)

# labeling each disorder based on classification
dt[, disorder_classification := ifelse(cause_id %in% mental_disorders, "mental",
                                       ifelse(cause_id %in% neuro_disorders, "neurological", NA))]

```

Adding super region and region name
```{r}
dt <- merge(dt, hierarchy[, .(location_id, super_region_name, region_name)], by = "location_id", all.x = T)

```

Adding interaction variable for cause and type of care
```{r}
dt[, causexcare_interaction := interaction(cause_name, type_of_care, sep = "_")]

```

Adding new age class columns to model dt
```{r}
# Age group class
dt[, age_group_class := fcase(
  grepl("under 1|0-14|5-11|6-18|5-14|8-17|7-18", population_age), "Children",
  grepl("12-18|12-17|<=17|<=18|15-19|15-17", population_age), "Adolescent",
  grepl("5-91|15\\+|18\\+|19\\+|20\\+|30\\+|40\\+|all ages", population_age), "All Ages",
  grepl("0-44|15-60|15-64|16-85|18-65|18-60|18-70|18-64|18-75|17\\+|20-60|20-64|20-68|19-78|20-60|20-64|16-75|under 50|under 65", population_age), "Adults",
  grepl("50\\+|55\\+|45-64|50-64", population_age), "Senior Adults",
  grepl("70\\+|65\\+|75\\+|80\\+|79-96|85-100|60\\+|65-75|74-88|65-79|over 60", population_age), "Elderly",
  default = "NO CLASS"
)]

dt[population_age == "under 5", age_group_class := "Children"]

# Population age adjusted
dt[, population_age_adj := fcase(
  grepl("8-17|12-17|12-18|15-17|15-19|6-18|7-18", population_age), "Teenagers",
  grepl("15-60|15-64|18-60|18-64|18-65|18-70|20-60|20-64|20-68|under 65", population_age), "Working Age",
  grepl("50-64|45-64", population_age), "Midlife",
  grepl("0-14|<=17|<=18", population_age), "<=18ish",
  grepl("15+|17+|18+|19+|20+|16-75|16-85|18-75|19-78", population_age), "18ish+",
  grepl("5-91|all ages", population_age), "All Ages",
  grepl("5-11|5-14", population_age), "Adolescence",
  grepl("0-44|under 50", population_age), "Under 50",
  grepl("over 60|60+", population_age), "60+",
  grepl("50+|55+", population_age), "50ish+",
  grepl("65-75|65-79|65+|70+|74-88|75+|79-96|80+|85-100", population_age), "Retired",
  default = "NO CLASS"
)]

dt[population_age_adj == "NO CLASS", population_age_adj := population_age]

```

Saving out dt 
```{r}
fwrite(dt, paste0(data_dir, "FILEPATH"))
```