Title:    05_generate_predicted_dataset.Rmd
 Project:  ##BHI - Brain Health Initiative
 Purpose : To create predicted dataset for regression
     
 Author: USERNAME
 Date: 2024 06 27
 
Setting up environment
```{r}
# clear environment
rm(list = ls())

# Set up directory roots #
if (!exists("code_repo"))  {
  code_repo <- unname(ifelse(Sys.info()['sysname'] == "Windows", "ADDRESS", paste0("ADDRESS", Sys.info()['user'][1], "ADDRESS")))
}

# packages
library(data.table)
library(lme4)
library(lmerTest)

# file paths
file_dir <- "FILEPATH"
direct_spending_path <- paste0(file_dir, "FILEPATH")
domestic_health_path <- "FILEPATH"
data_dir <- "FILEPATH"

# functions
source("FILEPATH")
source("FILEPATH")
source('FILEPATH')
source("FILEPATH")
source(paste0(code_repo, 'FILEPATH'))

# release number 
gbd_release_id <- 16

# get hierarchy
hierarchy <- get_location_metadata(
  location_set_id = 35,
  release_id = gbd_release_id
)

# storing country-level location ids
lvl3_locs <- sort(hierarchy[level == 3]$location_id)

# data info
years <- c(2000:2019)

tocs <- c("raked_mean_AM",
          "raked_mean_IP",
          "raked_mean_ED",
          "raked_mean_HH",
          "raked_mean_NF",
          "raked_mean_RX")

```

Import data
```{r}
# direct spending dataset
direct_spending_dt <- fread(direct_spending_path)

# domestic health spending dataset
dhe_dt <- fread(domestic_health_path)

```

Convert direct spending dataset from wide to long
```{r}
direct_spending_dt <- melt(direct_spending_dt, id.vars = c("location_id",
                                                   "location_name",
                                                   "year_id",
                                                   "age_group_id",
                                                   "age_group_name",
                                                   "sex_id",
                                                   "cause_id",
                                                   "cause_name",
                                                   "incident_prevalent_cases"), variable.name = "new_spending_type", value.name = "bhi_value",
                   measure.vars = c("AM", "ED", "HH", "IP", "NF", "RX"))

# adding type of care column
direct_spending_dt[, type_of_care := ifelse(new_spending_type == "AM", "raked_mean_AM",
                                                   ifelse(new_spending_type == "ED", "raked_mean_ED",
                                                          ifelse(new_spending_type == "HH", "raked_mean_HH",
                                                                 ifelse(new_spending_type == "IP", "raked_mean_IP",
                                                                        ifelse(new_spending_type == "NF", "raked_mean_NF",
                                                                               ifelse(new_spending_type == "RX", "raked_mean_RX", NA))))))]

# removing new spending type column
direct_spending_dt[, new_spending_type := NULL]

```

Aggregate data over ages and sex
```{r}
direct_spending_dt_agg <- data.table(aggregate(cbind(bhi_value, incident_prevalent_cases) ~ location_id + location_name + year_id + cause_id + cause_name + type_of_care, data = direct_spending_dt, FUN = sum))

# adding all age and all sex
direct_spending_dt_agg[, `:=` (age_group_id = 22,
                               sex_id = 3,
                               population_age = "all ages")]

```

Adding ihme loc id, super region, and super region name
```{r}
direct_spending_dt_agg <- merge(direct_spending_dt_agg, hierarchy[, .(location_id, ihme_loc_id, region_name, super_region_name)], by = "location_id", all.x = TRUE)

```

Creating spending per prevalent case values
```{r}
pred_dt <- copy(direct_spending_dt_agg)
pred_dt[, bhi_per_inc_prev_val := (bhi_value / incident_prevalent_cases)]

```

################################################################################
COVARIATES
################################################################################

GDP PC
```{r}
# get GDP per capita
gdp <- data.table(read_feather("FILEPATH"))
gdp <- gdp[scenario == 0 & year <= 2019 & year >= 2000]
gdp <- melt.data.table(gdp, id.vars = c("year", "iso3", "scenario"), value.name = "data_var", variable.name = "draw")
gdp <- gdp[, draw := str_replace(draw, "draw_", "")]
gdp <- gdp[, draw := as.numeric(as.character(draw))]
gdp[, scenario := NULL]

# condensing and renaming columns
gdp <- gdp[year %in% years,
           .(gdp_pc = mean(data_var)),
           by = .(year_id = year, ihme_loc_id = iso3)]

# creating gbd dataset
gdp <- merge(gdp[, .(ihme_loc_id, 
                        year_id, 
                        gdp_pc)], hierarchy[, .(location_id, 
                                                location_name, 
                                                ihme_loc_id)], by = 'ihme_loc_id')

# currency conversion
gdp_dt_conv <- currency_conversion(gdp,
                           col.loc = 'ihme_loc_id',
                           col.value = 'gdp_pc',
                           currency = 'usd',
                           currency.year = 2022 ,
                           base.year = 2021,
                           base.unit = 'usd',
                           converter.version = 7,
                           simplify = F)

# storing keep columns 
keep_cols <- c("ihme_loc_id", "year_id", "location_id", "currency_year_new", "gdp_pc_new")

# retaining keep columns
gdp_dt <- gdp_dt_conv[, ..keep_cols]

# renaming GDP per capita column
setnames(gdp_dt, old = "gdp_pc_new", new = "gdp_pc")

# removing extra currency year column
gdp_dt[, currency_year_new := NULL]

```

Fraction of population older than 65
```{r}
# getting total population values
tot_pops_dt <- get_population(location_id = lvl3_locs,
                              year_id = years,
                              age_group_id = 22,
                              sex_id = 3,
                              release_id = gbd_release_id 
                              )

# adding ihme loc id
tot_pops_dt <- merge(tot_pops_dt, hierarchy[, .(location_id, ihme_loc_id)], by = "location_id")

# condensing dataset
tot_pops_dt <- tot_pops_dt[, .(ihme_loc_id, 
                               location_id,
                               year_id, 
                               population)]

# get population values disaggregated by age
age_pop_dt <- get_population(location_id = lvl3_locs,
                       year_id = years,
                       age_group_id = c(1, 6:20, 30:32, 235),
                       sex_id = c(1, 2),
                       release_id = gbd_release_id 
                       )

# filter for population values of 65+
pop_over65 <- age_pop_dt[age_group_id %in% c(18:20, 30:32, 235), .(location_id, year_id, age_group_id, sex_id, population)]

# aggregate population values by country and year for population totals of 65+ 
pop_over65_agg <- data.table(aggregate(population ~ location_id + year_id, data = pop_over65, FUN = sum))

# rename population agg col
setnames(pop_over65_agg, old = "population", new = "pop_65_and_over")

# merge with total population values
pop_dt <- merge(pop_over65_agg, tot_pops_dt, by = c("location_id", "year_id"), all.x = T)

# calculate fraction of population that is 65+
pop_dt[, pop_65_frac := pop_65_and_over / population]

# remove extra total population column
pop_dt[, `:=` (population = NULL, pop_65_and_over = NULL)]

```

Adding socio-demographic index (SDI)
```{r}
# retrieving sdi values
sdi_dt <- get_covariate_estimates(covariate_id = 881,
                                   location_id = lvl3_locs,
                                   year_id =  years,
                                   release_id = gbd_release_id) 

# condensing dataset
sdi_dt <- sdi_dt[, .(location_id, 
                    year_id,
                    sdi_val = mean_value)]

```

Adding health access and quality index (HAQI)
```{r}
# retreiving haqi values
haqi <- get_covariate_estimates(covariate_id = 1099,
                                location_id = lvl3_locs,
                                year_id = years,
                                release_id = gbd_release_id)

# renaming haqi column
setnames(haqi, old = "mean_value", new = "haqi_val")

# condense dataset
haqi_dt <- haqi[, .(location_id, year_id, haqi_val)]

```

Adding domestic health spending data
```{r}
# retaining country-level data
dhe_dt <- dhe_dt[location_id %in% lvl3_locs]

```

Adding covariates to dataset
```{r}
pred_dt <- merge(pred_dt, gdp_dt, by = c("location_id", "ihme_loc_id", "year_id"), all.x = T)
pred_dt <- merge(pred_dt, pop_dt, by = c("location_id", "ihme_loc_id", "year_id"), all.x = T)
pred_dt <- merge(pred_dt, sdi_dt, by = c("location_id", "year_id"), all.x = T)
pred_dt <- merge(pred_dt, haqi_dt, by = c("location_id", "year_id"), all.x = T)
pred_dt <- merge(pred_dt, dhe_dt, by = c("location_id", "year_id"), all.x = T)

```

Adding sex label
```{r}
# adding sex label
pred_dt[sex_id == 3, sex_label := "Both Sexes"]

```

Adding disorder classification
```{r}
# mental brain health disorders
mental_disorders <- c(559,
                      560,
                      562,
                      566,
                    	567,
                    	570,
                    	571,
                    	572,
                    	575,
                    	578,
                      579,
                      582,
                      585,
                      718,
                      9999)

# neurological disorders
neuro_disorders <- c(332,
                     337,
                     477,
                     494,
                     543,
                     544,
                     545,
                     546,
                     554,
                     972)

# adding disorder classification
pred_dt[, disorder_classification := ifelse(cause_id %in% mental_disorders, "mental",
                                       ifelse(cause_id %in% neuro_disorders, "neurological", NA))]
```

Adding interaction variable for cause and type of care
```{r}
pred_dt[, causexcare_interaction := interaction(cause_name, type_of_care, sep = "_")]

```

Saving out prediction dataset
```{r}
fwrite(pred_dt, paste0(data_dir, "FILEPATH"))
```
