Title:    06a_regression_analysis.Rmd
 Project:  ##BHI - Brain Health Initiative
 Purpose : To create regression models for literature-based adjustment
     
 Author: USERNAME
 Date: 2025 03 07
 
Setting up environment
```{r}
# clear environment
rm(list = ls())

# Set up directory roots
if (!exists("code_repo"))  {
  code_repo <- unname(ifelse(Sys.info()['sysname'] == "Windows", "ADDRESS", paste0("ADDRESS", Sys.info()['user'][1], "ADDRESS")))
}

# packages
library(data.table)
library(lme4)
library(lmerTest)

# file paths
data_dir <- "FILEPATH"
model_data_path <- paste0(data_dir, "FILEPATH")
pred_shell_all_ages_agg_path <- paste0(data_dir, "FILEPATH")

# functions
source("FILEPATH")
source("FILEPATH")
source("FILEPATH")
source(paste0(code_repo, 'FILEPATH'))

# release number
gbd_release_id <- 16

# get hierarchy
hierarchy <- get_location_metadata(
  location_set_id = 35,
  release_id = gbd_release_id
)

```

Import data
```{r}
# model input data
model_dt <- fread(model_data_path)

# prediction dataset
pred_shell_all_ages_agg_dt <- fread(pred_shell_all_ages_agg_path)
pred_shell_all_ages_agg_dt[, age_group_class := "All Ages"]
pred_shell_all_ages_agg_dt[, population_age_adj := "All Ages"]

```

Data info
```{r}
locs <- sort(unique(model_dt$location_id))
years <- c(2000:2019)
age_groups <- c(1, 6:20, 30, 160)
sex <- c(1, 2)
```

Log transforming data
```{r}
# log transforming model input data
model_dt[, `:=` (lg_bhi_per_inc_prev_val = log(bhi_per_inc_prev_val), 
                 lg_lit_per_inc_prev_val = log(lit_per_inc_prev_val_new))]

# GDP per capita
model_dt[, lg_gdp_pc := log(gdp_pc)]

# Domestic health spending per capita
model_dt[, lg_dhe_pc := log(domestic_per_capita)]

# Government health spending per capita
model_dt[, lg_ghe_pc := log(ghes_per_capita)]

# Pulling all age and sex population estimates
model_locs <- sort(unique(model_dt$location_id))
model_pops <- get_population(location_id = model_locs, 
                       year_id = years, 
                       age_group_id = 22,
                       sex_id = 3,
                       release_id = gbd_release_id)

# Adding all age and sex population to model dataset
model_dt <- merge(model_dt, model_pops[, .(location_id,
                                           year_id,
                                           tote_pop = population)], by = c("location_id", "year_id"), all.x = T)

```

Log transforming covariates
```{r}
# setting offset value
bhi_offset_val <- 1e-10

# retaining pre-adjusted values
pred_shell_all_ages_agg_dt[, og_bhi_per_inc_prev_val := bhi_per_inc_prev_val]

# offsetting values
pred_shell_all_ages_agg_dt[, bhi_per_inc_prev_val := (bhi_per_inc_prev_val + bhi_offset_val)]

# log transforming bhi values
pred_shell_all_ages_agg_dt[, `:=` (lg_bhi_per_inc_prev_val = log(bhi_per_inc_prev_val), 
                                   lg_ghe_pc = log(ghes_per_capita))]
```

################################################################################
TYPE OF CARE SPECIFIC MODELS
################################################################################

Storing universal model info
```{r}
set.seed(123)

# set control
control_iter <- lmerControl(optimizer = "bobyqa",
                            optCtrl = list(maxfun = 200000),
                            check.conv.singular = .makeCC(action = "warning",
                                                          tol = 1e-6))
```

Identifying outliered data points
```{r}
# Creating proportional difference column
model_dt[, prop_diff := (lit_per_inc_prev_val_new / bhi_per_inc_prev_val)]

# setting threshold
threshold <- 20

# identifying outliered values
model_dt[, outlier_id := ifelse((prop_diff < (1/threshold) | prop_diff > (threshold)), 1, 0)]

```

AM Model
```{r}
# Filtering for toc
model_dt_AM <- model_dt[type_of_care == "raked_mean_AM" & outlier_id == 0]

# final model with regional intercepts
mod_AM <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_AM)

```

IP
```{r}
# Filtering for toc
model_dt_IP <- model_dt[type_of_care == "raked_mean_IP" & outlier_id == 0]

# final model with regional intercepts
mod_IP <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (1 | disorder_classification) +
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_IP)

```

RX
```{r}
# Filtering for toc
model_dt_RX <- model_dt[type_of_care == "raked_mean_RX" & outlier_id == 0]

# final model with regional intercepts
mod_RX <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_RX)

```

ED + HH + NF
```{r}
# Filtering for specific tocs
model_dt_ED_HH_NF <- model_dt[type_of_care %in% c("raked_mean_ED", 
                                                  "raked_mean_HH",
                                                  "raked_mean_NF") & outlier_id == 0]
# final model with regional and cause specific intercepts
mod_HH_ED_NF <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (0 + lg_bhi_per_inc_prev_val | location_name:type_of_care) +
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_ED_HH_NF)

```

Combining all model datasets
```{r}
model_input_dt <- rbindlist(list(model_dt_AM, model_dt_IP, model_dt_RX, model_dt_ED_HH_NF))

# removing extra columns
model_input_dt[, `:=` (mult_factor = NULL,
                       sample_size = NULL,
                       unit = NULL,
                       sampling_period = NULL,
                       base_year = NULL,
                       currency = NULL,
                       currency_col = NULL,
                       eur_usd = NULL,
                       deflator = NULL,
                       lcu_usd = NULL,
                       gdp_pc = NULL,
                       pop_65_frac = NULL,
                       population_age_adj = NULL,
                       haqi_val = NULL,
                       age_group_class = NULL)]
```


################################################################################
PREDICTED VALUES
################################################################################

################################################################################
AM
################################################################################

AM predict values
```{r}
# AM prediction dataset
pred_dt_AM <- pred_shell_all_ages_agg_dt[type_of_care == "raked_mean_AM"]


# predict function
pred_re_function_AM <- function(fit_mod){
                                predict(fit_mod, 
                                newdata = pred_dt_AM, 
                                allow.new.levels = TRUE, 
                                re.form = NULL) }

# Number of draws
sims <- 500
seed_val <- 123

# bootMer
boot_mod_AM <- bootMer(mod_AM, 
                       FUN = pred_re_function_AM,
                       type = "semiparametric",
                       nsim = sims,
                       seed = seed_val,
                       use.u = T)

# transpose bootMer outputs
boot_mod_AM_matrix <- t(boot_mod_AM[["t"]])

# convert to data.table
boot_mod_AM_draws_dt <- data.table(boot_mod_AM_matrix)

# draw columns
draw_col_names <- paste0("Draw_", 1:sims)
colnames(boot_mod_AM_draws_dt) <- draw_col_names

# exponentiate and remove offset
boot_mod_AM_draws_dt[, (draw_col_names) := lapply(.SD, function(x) exp(x)),.SDcols = draw_col_names]
boot_mod_AM_draws_dt[, (draw_col_names) := lapply(.SD, function(x) x - bhi_offset_val)]
boot_mod_AM_draws_dt[, (draw_col_names) := lapply(.SD, function(x) ifelse(x < 0, 0, x))]
                                
# adding ID row for merge
boot_mod_AM_draws_dt[, ID := 1:nrow(boot_mod_AM_draws_dt)]

# creating dataset shell for draws dataset
pred_dt_AM_draws <- pred_dt_AM[, .(location_id, 
                                   location_name, 
                                   year_id, 
                                   cause_name, 
                                   cause_id,
                                   type_of_care,
                                   disorder_classification,
                                   population_age,
                                   bhi_value,
                                   incident_prevalent_cases,
                                   og_bhi_per_inc_prev_val,
                                   bhi_per_inc_prev_val = (bhi_value / incident_prevalent_cases),
                                   ID = 1:nrow(boot_mod_AM_draws_dt))]

# merging draws dataset
pred_dt_AM_draws <- merge(pred_dt_AM_draws, boot_mod_AM_draws_dt, by = "ID", all.x = T)

# calculating upper and lower confidence intervals
pred_dt_AM_draws[, `:=` (draw_mean = rowMeans(.SD, na.rm = TRUE),
                         lwr = apply(.SD, 1, function(x) quantile(x, probs = 0.025, na.rm = TRUE)),
                         uppr = apply(.SD, 1, function(x) quantile(x, probs = 0.975, na.rm = TRUE))), .SDcols = draw_col_names]

```

################################################################################
IP
################################################################################
```{r}
# IP prediction dataset
pred_dt_IP <- pred_shell_all_ages_agg_dt[type_of_care == "raked_mean_IP"]

# predict function
pred_re_function_IP <- function(fit_mod){
                                predict(fit_mod, 
                                newdata = pred_dt_IP, 
                                allow.new.levels = TRUE, 
                                re.form = NULL) }

# number of draws
sims <- 500
seed_val <- 123

# bootMer
boot_mod_IP <- bootMer(mod_IP, 
                       FUN = pred_re_function_IP,
                       type = "semiparametric",
                       nsim = sims,
                       seed = seed_val,
                       use.u = T)

# transpose bootMer outputs
boot_mod_IP_matrix <- t(boot_mod_IP[["t"]])

# convert to data.table
boot_mod_IP_draws_dt <- data.table(boot_mod_IP_matrix)

# draw columns
draw_col_names <- paste0("Draw_", 1:sims)
colnames(boot_mod_IP_draws_dt) <- draw_col_names

# exponentiate and remove offset
boot_mod_IP_draws_dt[, (draw_col_names) := lapply(.SD, function(x) exp(x)),.SDcols = draw_col_names]
boot_mod_IP_draws_dt[, (draw_col_names) := lapply(.SD, function(x) x - bhi_offset_val)]
boot_mod_IP_draws_dt[, (draw_col_names) := lapply(.SD, function(x) ifelse(x < 0, 0, x))]
                                
# adding ID row for merge
boot_mod_IP_draws_dt[, ID := 1:nrow(boot_mod_IP_draws_dt)]

# creating dataset shell for draws dataset
pred_dt_IP_draws <- pred_dt_IP[, .(location_id, 
                                   location_name, 
                                   year_id, 
                                   cause_name, 
                                   cause_id,
                                   type_of_care,
                                   disorder_classification,
                                   population_age,
                                   bhi_value,
                                   incident_prevalent_cases,
                                   og_bhi_per_inc_prev_val,
                                   bhi_per_inc_prev_val = (bhi_value / incident_prevalent_cases),
                                   ID = 1:nrow(boot_mod_IP_draws_dt))]

# merging draws dataset
pred_dt_IP_draws <- merge(pred_dt_IP_draws, boot_mod_IP_draws_dt, by = "ID", all.x = T)

# calculating upper and lower confidence intervals
pred_dt_IP_draws[, `:=` (draw_mean = rowMeans(.SD, na.rm = TRUE),
                         lwr = apply(.SD, 1, function(x) quantile(x, probs = 0.025, na.rm = TRUE)),
                         uppr = apply(.SD, 1, function(x) quantile(x, probs = 0.975, na.rm = TRUE))), .SDcols = draw_col_names]
```


################################################################################
RX
################################################################################

RX Predicted values
```{r}
# RX prediction dataset
pred_dt_RX <- pred_shell_all_ages_agg_dt[type_of_care == "raked_mean_RX"]

# predict function
pred_re_function_RX <- function(fit_mod){
                                predict(fit_mod, 
                                newdata = pred_dt_RX, 
                                allow.new.levels = TRUE, 
                                re.form = NULL) }

# number of draws
sims <- 500
seed_val <- 123

# bootMer
boot_mod_RX <- bootMer(mod_RX, 
                       FUN = pred_re_function_RX,
                       type = "semiparametric",
                       nsim = sims,
                       seed = seed_val,
                       use.u = T)

# transpose bootMer outputs
boot_mod_RX_matrix <- t(boot_mod_RX[["t"]])

# convert to data.table
boot_mod_RX_draws_dt <- data.table(boot_mod_RX_matrix)

# draw columns
draw_col_names <- paste0("Draw_", 1:sims)
colnames(boot_mod_RX_draws_dt) <- draw_col_names

# exponentiate and remove offset
boot_mod_RX_draws_dt[, (draw_col_names) := lapply(.SD, function(x) exp(x)),.SDcols = draw_col_names]
boot_mod_RX_draws_dt[, (draw_col_names) := lapply(.SD, function(x) x - bhi_offset_val)]
boot_mod_RX_draws_dt[, (draw_col_names) := lapply(.SD, function(x) ifelse(x < 0, 0, x))]
                                
# adding ID row for merge
boot_mod_RX_draws_dt[, ID := 1:nrow(boot_mod_RX_draws_dt)]

# creating dataset shell for draws dataset
pred_dt_RX_draws <- pred_dt_RX[, .(location_id, 
                                   location_name, 
                                   year_id, 
                                   cause_name, 
                                   cause_id,
                                   type_of_care,
                                   disorder_classification,
                                   population_age,
                                   bhi_value,
                                   incident_prevalent_cases,
                                   og_bhi_per_inc_prev_val,
                                   bhi_per_inc_prev_val = (bhi_value / incident_prevalent_cases),
                                   ID = 1:nrow(boot_mod_RX_draws_dt))]

# merging draws dataset
pred_dt_RX_draws <- merge(pred_dt_RX_draws, boot_mod_RX_draws_dt, by = "ID", all.x = T)

# calculating upper and lower confidence intervals
pred_dt_RX_draws[, `:=` (draw_mean = rowMeans(.SD, na.rm = TRUE),
                         lwr = apply(.SD, 1, function(x) quantile(x, probs = 0.025, na.rm = TRUE)),
                         uppr = apply(.SD, 1, function(x) quantile(x, probs = 0.975, na.rm = TRUE))), .SDcols = draw_col_names]
```

################################################################################
ED + HH + NF
################################################################################

```{r}
# ED + HH + NF prediction dataset
pred_dt_ED_HH_NF <- pred_shell_all_ages_agg_dt[type_of_care %in% c("raked_mean_ED",
                                                                   "raked_mean_HH",
                                                                   "raked_mean_NF")]

# predict function
pred_re_function_ED_HH_NF <- function(fit_mod){
                                predict(fit_mod, 
                                newdata = pred_dt_ED_HH_NF, 
                                allow.new.levels = TRUE, 
                                re.form = NULL) }

# number of draws
sims <- 500
seed_val <- 123

# bootMer
boot_mod_ED_HH_NF <- bootMer(mod_HH_ED_NF, 
                             FUN = pred_re_function_ED_HH_NF,
                             type = "semiparametric",
                             nsim = sims,
                             seed = seed_val,
                             use.u = T)

# transpose bootMer outputs
boot_mod_ED_HH_NF_matrix <- t(boot_mod_ED_HH_NF[["t"]])

# convert to data.table
boot_mod_ED_HH_NF_draws_dt <- data.table(boot_mod_ED_HH_NF_matrix)

# draw columns
draw_col_names <- paste0("Draw_", 1:sims)
colnames(boot_mod_ED_HH_NF_draws_dt) <- draw_col_names

# exponentiate and remove offset
boot_mod_ED_HH_NF_draws_dt[, (draw_col_names) := lapply(.SD, function(x) exp(x)),.SDcols = draw_col_names]
boot_mod_ED_HH_NF_draws_dt[, (draw_col_names) := lapply(.SD, function(x) x - bhi_offset_val)]
boot_mod_ED_HH_NF_draws_dt[, (draw_col_names) := lapply(.SD, function(x) ifelse(x < 0, 0, x))]
                                
# adding ID row for merge
boot_mod_ED_HH_NF_draws_dt[, ID := 1:nrow(boot_mod_ED_HH_NF_draws_dt)]

# creating dataset shell for draws dataset
pred_dt_ED_HH_NF_draws <- pred_dt_ED_HH_NF[, .(location_id, 
                                               location_name, 
                                               year_id, 
                                               cause_name, 
                                               cause_id,
                                               type_of_care,
                                               disorder_classification,
                                               population_age,
                                               bhi_value,
                                               incident_prevalent_cases,
                                               og_bhi_per_inc_prev_val,
                                               bhi_per_inc_prev_val = (bhi_value / incident_prevalent_cases),
                                               ID = 1:nrow(boot_mod_ED_HH_NF_draws_dt))]

# merging draws dataset
pred_dt_ED_HH_NF_draws <- merge(pred_dt_ED_HH_NF_draws, boot_mod_ED_HH_NF_draws_dt, by = "ID", all.x = T)

# calculating upper and lower confidence intervals
pred_dt_ED_HH_NF_draws[, `:=` (draw_mean = rowMeans(.SD, na.rm = TRUE),
                         lwr = apply(.SD, 1, function(x) quantile(x, probs = 0.025, na.rm = TRUE)),
                         uppr = apply(.SD, 1, function(x) quantile(x, probs = 0.975, na.rm = TRUE))), .SDcols = draw_col_names]
```

Combining all type of care predicted values datasets
```{r}
pred_dt_draws <- rbindlist(list(pred_dt_AM_draws, pred_dt_IP_draws, pred_dt_RX_draws, pred_dt_ED_HH_NF_draws))
```

Condensing predicted draws dataset
```{r}
pred_dt_draws_cond <- pred_dt_draws[, .(location_id, 
                                        location_name, 
                                        year_id, 
                                        cause_name, 
                                        cause_id,
                                        type_of_care,
                                        disorder_classification,
                                        population_age,
                                        bhi_value,
                                        incident_prevalent_cases,
                                        bhi_per_inc_prev_val,
                                        og_bhi_per_inc_prev_val,
                                        draw_mean,
                                        lwr,
                                        uppr)]

```

Saving out modeled dataset
```{r}
fwrite(pred_dt_draws, "FILEPATH")

```

Saving out modeled input dataset
```{r}
fwrite(model_dt, "FILEPATH")

```