Title:    06b_sensitivity_analysis.Rmd
 Project:  ##BHI - Brain Health Initiative
 Purpose : Sensitivity analysis for literature-based adjustment excluding data from 2016 to 2019
     
 Author: USERNAME
 Date: 2025 01 22
 
Setting up environment
```{r}
# clear environment
rm(list = ls())

# Set up directory roots #
if (!exists("code_repo"))  {
  code_repo <- unname(ifelse(Sys.info()['sysname'] == "Windows", "ADDRESS", paste0("ADDRESS", Sys.info()['user'][1], "ADDRESS")))
}

# packages
library(data.table)
library(lme4)
library(lmerTest)

# file paths
data_dir <- "FILEPATH"
data_path <- paste0(data_dir, "FILEPATH")
model_outputs_dir <- paste0(data_dir, "FILEPATH")
pred_shell_all_ages_agg_path <- paste0(data_dir, "FILEPATH")
fig_out_dir <- paste0(data_dir, "FILEPATH")

# functions
source("FILEPATH")
source("FILEPATH")
source("FILEPATH")

# release id
gbd_release_id <- 16

# get hierarchy
hierarchy <- get_location_metadata(
  location_set_id = 35,
  release_id = gbd_release_id
)

```

Import data
```{r}
# model input data
model_dt <- fread(data_path)

# predicted dataset
pred_shell_all_ages_agg_dt <- fread(pred_shell_all_ages_agg_path)
pred_shell_all_ages_agg_dt[, age_group_class := "All Ages"]
pred_shell_all_ages_agg_dt[, population_age_adj := "All Ages"]

```

Data info
```{r}
locs <- sort(unique(model_dt$location_id))
years <- c(2000:2019)
age_groups <- c(1, 6:20, 30, 160)
sex <- c(1, 2)
```

Log transforming data
```{r}
# log transforming model input data
model_dt[, `:=` (lg_bhi_per_inc_prev_val = log(bhi_per_inc_prev_val), 
                 lg_lit_per_inc_prev_val = log(lit_per_inc_prev_val_new))]

# GDP per capita
model_dt[, lg_gdp_pc := log(gdp_pc)]

# Domestic health spending per capita
model_dt[, lg_dhe_pc := log(domestic_per_capita)]

# Government health spending per capita
model_dt[, lg_ghe_pc := log(ghes_per_capita)]

# Pulling all age and all sex population estimates
model_locs <- sort(unique(model_dt$location_id))
model_pops <- get_population(location_id = model_locs, 
                       year_id = years, 
                       age_group_id = 22,
                       sex_id = 3,
                       release_id = gbd_release_id)

# adding all age and all sex population to model dataset
model_dt <- merge(model_dt, model_pops[, .(location_id,
                                           year_id,
                                           tote_pop = population)], by = c("location_id", "year_id"), all.x = T)

```

Log transforming prediction shell dataset covariates
```{r}
# set offset
bhi_offset_val <- 1e-10

# retaining pre-adjusted bhi values
pred_shell_all_ages_agg_dt[, og_bhi_per_inc_prev_val := bhi_per_inc_prev_val]

# adding offset
pred_shell_all_ages_agg_dt[, bhi_per_inc_prev_val := (bhi_per_inc_prev_val + bhi_offset_val)]

# log transforming bhi values
pred_shell_all_ages_agg_dt[, `:=` (lg_bhi_per_inc_prev_val = log(bhi_per_inc_prev_val), 
                                   lg_ghe_pc = log(ghes_per_capita))]
```

################################################################################
TYPE OF CARE SPECIFIC MODELS
################################################################################

Storing universal model info
```{r}
set.seed(123)

# set controls
control_iter <- lmerControl(optimizer = "bobyqa",
                            optCtrl = list(maxfun = 200000),
                            check.conv.singular = .makeCC(action = "warning",
                                                          tol = 1e-6))
```

Identifying outliered data points
```{r}
# Creating proportional difference column
model_dt[, prop_diff := (lit_per_inc_prev_val_new / bhi_per_inc_prev_val)]

# set threshold
threshold <- 20

# identifying outliered values
model_dt[, outlier_id := ifelse((prop_diff < (1/threshold) | prop_diff > (threshold)), 1, 0)]

```

AM Model
```{r}
# Filtering for toc
model_dt_AM <- model_dt[type_of_care == "raked_mean_AM" & outlier_id == 0]

# Final model with regional intercepts
mod_AM <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_AM)

```

Exclude data from 2016 to 2019
```{r}
model_dt_AM2 <- model_dt_AM[!year_id %in% c(2016:2019)]
```

AM Model #2 
```{r}
# ambulatory sensitivity model
mod_AM2 <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_AM2)

```

AM Predicted Values 
```{r}
model_dt_AM_sa <- copy(model_dt_AM)

# original model
model_dt_AM_sa[, `:=` (pred_mod_am_vals = predict(mod_AM, newdata = model_dt_AM_sa, allow.new.levels = TRUE),
                       pred_mod_am2_vals = predict(mod_AM2, newdata = model_dt_AM_sa, allow.new.levels = TRUE))]

# exponentiate predicted values
model_dt_AM_sa[, `:=` (pred_mod_am_vals = exp(pred_mod_am_vals),
                       pred_mod_am2_vals = exp(pred_mod_am2_vals))]
```

################################################################################
AM TABLE
################################################################################
```{r}
AM_table <- model_dt_AM_sa[year_id %in% c(2016:2019), .(mae = mean(abs(lit_per_inc_prev_val_new - pred_mod_am_vals)),
                                   mape = mean(abs(lit_per_inc_prev_val_new - pred_mod_am2_vals)),
                                   mse = mean((lit_per_inc_prev_val_new - pred_mod_am_vals)^2),
                                   mspe = mean((lit_per_inc_prev_val_new - pred_mod_am2_vals)^2)), by = .(year_id)]

```


IP
```{r}
# Filtering for toc
model_dt_IP <- model_dt[type_of_care == "raked_mean_IP" & outlier_id == 0]

# final model with regional intercepts
mod_IP <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (1 | disorder_classification) + 
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_IP)

```

Exclude data from 2016 to 2019
```{r}
model_dt_IP2 <- model_dt_IP[!year_id %in% c(2016:2019)]
```

IP Model #2
```{r}
# inpatient sensitivity model
mod_IP2 <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (1 | disorder_classification) + 
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_IP2)

```

IP Predicted Values 
```{r}
model_dt_IP_sa <- copy(model_dt_IP)

# Creating predicted values
model_dt_IP_sa[, pred_mod_IP_vals := predict(mod_IP, newdata = model_dt_IP_sa, allow.new.levels = TRUE)]
model_dt_IP_sa[, pred_mod_IP2_vals := predict(mod_IP2, newdata = model_dt_IP_sa, allow.new.levels = TRUE)]

# Exponentiate values
model_dt_IP_sa[, `:=` (pred_mod_IP_vals = exp(pred_mod_IP_vals),
                       pred_mod_IP2_vals = exp(pred_mod_IP2_vals))]
```

################################################################################
IP TABLE
################################################################################
```{r}
IP_table <- model_dt_IP_sa[year_id %in% c(2016:2019), .(mae = mean(abs(lit_per_inc_prev_val_new - pred_mod_IP_vals)),
                                   mape = mean(abs(lit_per_inc_prev_val_new - pred_mod_IP2_vals)),
                                   mse = mean((lit_per_inc_prev_val_new - pred_mod_IP_vals)^2),
                                   mspe = mean((lit_per_inc_prev_val_new - pred_mod_IP2_vals)^2)), by = .(year_id)]
```

################################################################################
RX TABLE
################################################################################

RX
```{r}
# Filtering for toc
model_dt_RX <- model_dt[type_of_care == "raked_mean_RX" & outlier_id == 0]

# final model with regional intercepts
mod_RX <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_RX)

```

Exclude data from 2016 to 2019
```{r}
model_dt_RX2 <- model_dt_RX[!year_id %in% c(2016:2019)]
```

RX Model #2 
```{r}
# RX sensitivity model
mod_RX2 <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_RX2)

```

RX Predicted Values 
```{r}
model_dt_RX_sa <- copy(model_dt_RX)

# Creating predicted values
model_dt_RX_sa[, pred_mod_RX_vals := predict(mod_RX, newdata = model_dt_RX_sa, allow.new.levels = TRUE)]
model_dt_RX_sa[, pred_mod_RX2_vals := predict(mod_RX2, newdata = model_dt_RX_sa, allow.new.levels = TRUE)]

# Exponentiate values
model_dt_RX_sa[, `:=` (pred_mod_RX_vals = exp(pred_mod_RX_vals),
                       pred_mod_RX2_vals = exp(pred_mod_RX2_vals))]

```

################################################################################
RX TABLE
################################################################################
```{r}
RX_table <- model_dt_RX_sa[year_id %in% c(2016:2019), .(mae = mean(abs(lit_per_inc_prev_val_new - pred_mod_RX_vals)),
                                   mape = mean(abs(lit_per_inc_prev_val_new - pred_mod_RX2_vals)),
                                   mse = mean((lit_per_inc_prev_val_new - pred_mod_RX_vals)^2),
                                   mspe = mean((lit_per_inc_prev_val_new - pred_mod_RX2_vals)^2)), by = .(year_id)]
```


################################################################################

ED + HH + NF
```{r}
# Filtering for specific tocs
model_dt_ED_HH_NF <- model_dt[type_of_care %in% c("raked_mean_ED", 
                                                  "raked_mean_HH",
                                                  "raked_mean_NF") & outlier_id == 0]

# final model with regional and cause-specific intercepts
mod_ED_HH_NF <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (0 + lg_bhi_per_inc_prev_val | location_name:type_of_care) +
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_ED_HH_NF)

```

Exclude data from 2016 to 2019
```{r}
model_dt_ED_HH_NF2 <- model_dt_ED_HH_NF[!year_id %in% c(2016:2019)]
```

ED + HH + NF Model #2
```{r}
# group sensitivity model
mod_ED_HH_NF2 <- lmer(lg_lit_per_inc_prev_val ~ -1 + lg_bhi_per_inc_prev_val +
                        (0 + lg_bhi_per_inc_prev_val | location_name:type_of_care) +
                        (0 + lg_bhi_per_inc_prev_val | super_region_name/region_name:cause_name) +
                        (1 + lg_bhi_per_inc_prev_val | location_name:cause_name),
    	                  REML = FALSE,
    	                  control = control_iter,
    	                  data = model_dt_ED_HH_NF2)

```

Group model predicted values 
```{r}
model_dt_ED_HH_NF_sa <- copy(model_dt_ED_HH_NF)

# Creating predicted values
model_dt_ED_HH_NF_sa[, pred_mod_ED_HH_NF_vals := predict(mod_ED_HH_NF, newdata = model_dt_ED_HH_NF_sa, allow.new.levels = TRUE)]
model_dt_ED_HH_NF_sa[, pred_mod_ED_HH_NF2_vals := predict(mod_ED_HH_NF2, newdata = model_dt_ED_HH_NF_sa, allow.new.levels = TRUE)]

# Exponentiate values
model_dt_ED_HH_NF_sa[, `:=`(pred_mod_ED_HH_NF_vals = exp(pred_mod_ED_HH_NF_vals),
                       pred_mod_ED_HH_NF2_vals = exp(pred_mod_ED_HH_NF2_vals))]
```

################################################################################
ED HH NF Table
################################################################################
```{r}
ED_HH_NF_table <- model_dt_ED_HH_NF_sa[year_id %in% c(2016:2019), .(mae = mean(abs(lit_per_inc_prev_val_new - pred_mod_ED_HH_NF_vals)),
                                   mape = mean(abs(lit_per_inc_prev_val_new - pred_mod_ED_HH_NF2_vals)),
                                   mse = mean((lit_per_inc_prev_val_new - pred_mod_ED_HH_NF_vals)^2),
                                   mspe = mean((lit_per_inc_prev_val_new - pred_mod_ED_HH_NF2_vals)^2)), by = .(year_id)]
```

Creating MAE consolidated table
```{r}
AM_table[, Model := "Ambulatory"]
IP_table[, Model := "Inpatient"]
RX_table[, Model := "Pharmaceuticals"]
ED_HH_NF_table[, Model := "Combine model (ED, HH, NF)"]

# append tables
sensitivity_analysis_table <- rbindlist(list(AM_table, IP_table, RX_table, ED_HH_NF_table))

# reorder table columns
sensitivity_analysis_table <- sensitivity_analysis_table[, .(Model, 
                         year_id,
                         mae,
                         mape,
                         mse,
                         mspe)]

# renaming measure columns
setnames(sensitivity_analysis_table, c("mae", "mape", "mse", "mspe", "year_id"), 
         c("MAE Full Model", "MAPE OOS", "MSE Full Model", "MSPE OOS", "Year"))

sensitivity_analysis_table <- sensitivity_analysis_table[order(Year, Model)]

# round values in mean sensitivity analysis table
sensitivity_analysis_table[, `:=` (`MAE Full Model` = round(`MAE Full Model`, 0),
                 `MAPE OOS` = round(`MAPE OOS`, 0),
                 `MSE Full Model` = round(`MSE Full Model`, 0),
                 `MSPE OOS` = round(`MSPE OOS`, 0))]
```

Saving out MAE table and excluded data
```{r}
fwrite(sensitivity_analysis_table, "FILEPATH")

```